---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# voteReproject

<!-- badges: start -->
<!-- badges: end -->

The 2017, 2019 and 2024 UK General Elections produced highly contrasting results. 

It is, however, difficult to compare certain aspects of them because of boundary changes to constituencies which took effect for the 2024 election. 

The goal of `voteReproject` is, by reprojecting voting data from the earlier elections to the 2024 boundaries, to provide datasets which are suitable for comparative analysis.

## Installation

You can install the development version of voteReproject from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("horankev/voteReproject")
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(voteReproject)
library(tidyverse)
library(sf)
library(ggpubr)
library(here)
library(kableExtra)

```

## Boundary changes

The 2024 and 2017/19 constituency boundaries for UK General Elections are mapped below.

```{r, fig.width=10, fig.height=10, echo=FALSE}

temp <- data.frame(
  geometry = c(votes19$geometry, votes24$geometry),
  year = c(rep("2019",650),rep("2024",650))
) |> 
  st_as_sf()

ggplot(temp) +
  geom_sf(aes(colour=year, linewidth=year), fill=NA) +
  scale_colour_manual(values = c("#ED1941","gray50")) +
  scale_linewidth_manual(values = c(0.2,0.5)) +
  theme_void()

```

## Reprojection

Votes are allocated to new (2024) constituencies from the boundaries which existed in 2017/19 according to the proportion of the population living in areas of overlap. The first five entries of the table used for this projection are shown below.

```{r, echo=FALSE}

all_changes <- readxl::read_xlsx(here("data-raw","boundaries","Boundary_changes_data_file.xlsx"), 
                                 sheet = 3, skip = 1)

projection_df <- all_changes |> 
  select(1:4,9,11) |> 
  rename(prop_old_in_new = `Percentage of old constituency in this segment (population) [notes 1 and 2]`) |> 
  # make alterations for Scottish constituencies with different codes in 2024
  # none of these constituencies have had their boundaries moved
  # just changing code manually
  mutate(`New constituency code` = case_when(`New constituency code` == "S14000006" ~ "S14000107",
                                             `New constituency code` == "S14000008" ~ "S14000108",
                                             `New constituency code` == "S14000010" ~ "S14000109",
                                             `New constituency code` == "S14000040" ~ "S14000110",
                                             `New constituency code` == "S14000058" ~ "S14000111",
                                             TRUE ~ `New constituency code`)) |> 
  mutate(`Current constituency code` = case_when(`Current constituency code` == "S14000006" ~ "S14000107",
                                                 `Current constituency code` == "S14000008" ~ "S14000108",
                                                 `Current constituency code` == "S14000010" ~ "S14000109",
                                                 `Current constituency code` == "S14000040" ~ "S14000110",
                                                 `Current constituency code` == "S14000058" ~ "S14000111",
                                                 TRUE ~ `Current constituency code`))

kable(head(projection_df, 5), booktabs = TRUE, caption = "source: <https://commonslibrary.parliament.uk/boundary-review-2023-which-seats-will-change/>")
```

## Example

Below are shown the winning party in each constituency from the 2024 election and the first-placed party from the two previous elections subject to reprojection. The boundaries are sourced from <https://geoportal.statistics.gov.uk> and the voting data from <https://commonslibrary.parliament.uk/tag/elections-data/>.

```{r example1, fig.width=12, fig.height=4}

ggarrange(
  ggplot(votestogether) +
    geom_sf(aes(fill=first_party24)) +
    scale_fill_manual(values = party_palette) + 
    theme_void(),
  
  ggplot(votestogether) +
    geom_sf(aes(fill=first_party19)) +
    scale_fill_manual(values = party_palette) + 
    theme_void(),
  
  ggplot(votestogether) +
    geom_sf(aes(fill=first_party17)) +
    scale_fill_manual(values = party_palette) + 
    theme_void(),
  
  ncol = 3
)


```

The problem of invisibility of constituencies which are small by area, usually dense urban settings, can be overcome by using hexagonal representations of the constituencies. Their positioning aims to closely resemble their actual relative position. Here, we use hexagons created by Philip Brown and Alasdair Rae from Automatic Knowledge (<https://observablehq.com/@jwolondon/uk-election-2024-boundary-data>).

```{r example2, fig.width=12, fig.height=4}

ggarrange(
  ggplot(votestogether_hex) +
    geom_sf(aes(fill=first_party24)) +
    scale_fill_manual(values = party_palette) + 
    theme_void(),
  
  ggplot(votestogether_hex) +
    geom_sf(aes(fill=first_party19)) +
    scale_fill_manual(values = party_palette) + 
    theme_void(),
  
  ggplot(votestogether_hex) +
    geom_sf(aes(fill=first_party17)) +
    scale_fill_manual(values = party_palette) + 
    theme_void(),
  
  ncol = 3
)


```


Looking at changes in votes in Scotland for the Scottish National Party and Labour: 

```{r, fig.width=8, fig.height=5}

ggarrange(
  ggplot(votestogether |> filter(region_name == "Scotland")) +
    geom_sf(aes(fill=snp24_pct)) +
    scale_fill_distiller(palette = "Oranges", direction = 1, limits = c(0,55)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(region_name == "Scotland")) +
    geom_sf(aes(fill=snp19_pct)) +
    scale_fill_distiller(palette = "Oranges", direction = 1, limits = c(0,55)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(region_name == "Scotland")) +
    geom_sf(aes(fill=snp17_pct)) +
    scale_fill_distiller(palette = "Oranges", direction = 1, limits = c(0,55)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(region_name == "Scotland")) +
    geom_sf(aes(fill=lab24_pct)) +
    scale_fill_distiller(palette = "Reds", direction = 1, limits = c(0,60)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(region_name == "Scotland")) +
    geom_sf(aes(fill=lab19_pct)) +
    scale_fill_distiller(palette = "Reds", direction = 1, limits = c(0,60)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(region_name == "Scotland")) +
    geom_sf(aes(fill=lab17_pct)) +
    scale_fill_distiller(palette = "Reds", direction = 1, limits = c(0,60)) + 
    theme_void(),
  
  ncol = 3,
  nrow = 2
)

```

Looking at changes in votes in England for the Conservatives and Labour: 

```{r, fig.width=10, fig.height=5}

ggarrange(
  ggplot(votestogether |> filter(!region_name %in% c("Wales","Scotland","Northern Ireland"))) +
    geom_sf(aes(fill=con24_pct, colour=con24_pct)) +
    scale_fill_distiller(palette = "Blues", direction = 1, limits = c(0,80)) + 
    scale_colour_distiller(palette = "Blues", direction = 1, limits = c(0,80)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(!region_name %in% c("Wales","Scotland","Northern Ireland"))) +
    geom_sf(aes(fill=con19_pct, colour=con19_pct)) +
    scale_fill_distiller(palette = "Blues", direction = 1, limits = c(0,80)) + 
    scale_colour_distiller(palette = "Blues", direction = 1, limits = c(0,80)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(!region_name %in% c("Wales","Scotland","Northern Ireland"))) +
    geom_sf(aes(fill=con17_pct, colour=con17_pct)) +
    scale_fill_distiller(palette = "Blues", direction = 1, limits = c(0,80)) + 
    scale_colour_distiller(palette = "Blues", direction = 1, limits = c(0,80)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(!region_name %in% c("Wales","Scotland","Northern Ireland"))) +
    geom_sf(aes(fill=lab24_pct, colour=lab24_pct)) +
    scale_fill_distiller(palette = "Reds", direction = 1, limits = c(0,80)) + 
    scale_colour_distiller(palette = "Reds", direction = 1, limits = c(0,80)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(!region_name %in% c("Wales","Scotland","Northern Ireland"))) +
    geom_sf(aes(fill=lab19_pct, colour=lab19_pct)) +
    scale_fill_distiller(palette = "Reds", direction = 1, limits = c(0,80)) + 
    scale_colour_distiller(palette = "Reds", direction = 1, limits = c(0,80)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(!region_name %in% c("Wales","Scotland","Northern Ireland"))) +
    geom_sf(aes(fill=lab17_pct, colour=lab17_pct)) +
    scale_fill_distiller(palette = "Reds", direction = 1, limits = c(0,80)) + 
    scale_colour_distiller(palette = "Reds", direction = 1, limits = c(0,80)) + 
    theme_void(),
  
  ncol = 3,
  nrow = 2
)

```

Looking at changes in votes in Northern Ireland for Sinn Fein and the Democratic Unionist Party: 

```{r, fig.width=12, fig.height=5}

ggarrange(
  ggplot(votestogether |> filter(region_name == "Northern Ireland")) +
    geom_sf(aes(fill=sf24_pct)) +
    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,70)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(region_name == "Northern Ireland")) +
    geom_sf(aes(fill=sf19_pct)) +
    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,70)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(region_name == "Northern Ireland")) +
    geom_sf(aes(fill=sf17_pct)) +
    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,70)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(region_name == "Northern Ireland")) +
    geom_sf(aes(fill=dup24_pct)) +
    scale_fill_distiller(palette = "Blues", direction = 1, limits = c(0,70)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(region_name == "Northern Ireland")) +
    geom_sf(aes(fill=dup19_pct)) +
    scale_fill_distiller(palette = "Blues", direction = 1, limits = c(0,70)) + 
    theme_void(),
  
  ggplot(votestogether |> filter(region_name == "Northern Ireland")) +
    geom_sf(aes(fill=dup17_pct)) +
    scale_fill_distiller(palette = "Blues", direction = 1, limits = c(0,70)) + 
    theme_void(),
  
  ncol = 3,
  nrow = 2
)

```





